---
title: "homework_6"
author: "Harinath Reddy"
date: "2022-11-24"
output:
  pdf_document: default
  html_document: default
---



```{r}
library(dplyr)
library(tsibble)
library(fable)
library(feasts)
library(tidyr)
library(ggplot2)
```

```{r}

library(IDDA)
data(state.ts)
```


```{r}
## Using `DATE` as index variable.
Virginia.ts <- state.ts %>%
  dplyr::filter(State == "Virginia") %>%
  dplyr::select(Death, Y.Death) %>%
  mutate(YDA_Death = lag(Y.Death)) %>%
  dplyr::filter(!is.na(YDA_Death))

head(Virginia.ts )
```

## Creatiung the Training Data
```{r}

# time series plot death counts in Florida
Virginia.ts %>%
  autoplot(Y.Death)
```


# set training data from NOV 28 to DEC 04
```{r}
train <- Virginia.ts %>% 
  filter_index("2020-01-23" ~ "2020-12-17
               ")
n <- nrow(train)
```



```{r}
fit_trends <- train %>%
  model(
    Linear = TSLM(Y.Death ~ trend()),
    exponential = TSLM(log(Y.Death + 1) ~ trend()),
  )

fc_trends <- fit_trends %>% fabletools::forecast( h = 7)

```


## Making Predictions 
```{r}
Virginia.ts %>%
  dplyr::filter(DATE < train$DATE[n] + 7 ) %>%
  autoplot(Y.Death) +
  geom_line(data = fitted(fit_trends),
            aes(y = .fitted, color = .model)) +
  autolayer(fc_trends, alpha = 0.5, level = 95) +
  labs(y = "Number of deaths",
       title = "Daily new deaths in Virginia") +
  theme_bw() +
  theme(legend.position = "bottom")

```

```{r}

lm_fit <- train%>%
  model(lm = TSLM(Y.Death ~ log(YDA_Death + 1)))
report(lm_fit)


```


## Fitting the ETS Model
```{r}
ets_fit <-train %>%
  model(ETS(Y.Death ~ error("A") + trend("A") + season("A"), opt_crit =
"mse"))
report(ets_fit)
```

```{r}

residual_plot_linear_model <- gg_tsresiduals(lm_fit)

residual_plot_linear_model
```

```{r}
residual_plot_ets_model <- gg_tsresiduals(ets_fit)

residual_plot_ets_model


```


## Yes the residuals appear to be reasonably normally distributed



## Piecewise constant spline regression model with 15 interior knots
```{r}

n <- nrow(Virginia.ts)
t <- 1:n
y <- Virginia.ts$Y.Death

# Knots
N <- 15
knots <- 1 + (n-1)/(N+1) * (0:N)

# Piecewise constant spline basis
t.rep <- matrix(rep(t, N), n, N)
knot.L <- matrix(rep(knots[-(N + 1)], each = n), n, N)
knot.R <- matrix(rep(knots[-1], each = n), n, N)
B <- 1*((knot.L <= t.rep) & (t.rep < knot.R))
X <- cbind(B, knots[N] < t & t <= n)

M <- t(X) %*% X
beta <- solve(M) %*% t(X) %*% y
yhat <- X %*% beta
Virginia.ts$pcs_preds <- yhat
```



```{r}

# plot of reported vs piecewise constant spline fit
pcs_p <- Virginia.ts %>%
  ggplot(aes(x = DATE)) +
  geom_line(aes(y = Y.Death, color = "Reported")) +
  geom_line(aes(y = pcs_preds, color = "Fitted")) +
  scale_color_manual(
    values = c(Reported = "black", Fitted = "red")
  ) +
  labs(y = "Death count",
       title = "Reported vs piecewise constant spline fit") +
  guides(color = guide_legend(title = "Series"))

pcs_p


```

### Truncated Power Spline 

```{r}

y <- Virginia.ts$Y.Death
n <- nrow(Virginia.ts)
t <- 1:n

# knots
N <- 10
knots <- 1 + (n-1)/(N+1) * (0:N)

# truncated power spline basis functions
X <- matrix(1, n, N + 2)	
X[, 2] <- t
t.rep <- matrix(rep(t, N), n, N) 
tmp <- t.rep - matrix(rep(knots[2:(N + 1)], each = n), n, N)
X[, 3:(N+2)] <- tmp * (tmp > 0)

# truncated power spline fit
M <- t(X) %*% X
beta <- solve(M) %*% t(X) %*% y
yhat <- X %*% beta
Virginia.ts$tps_preds <- yhat

# plot of reported vs truncated power spline fit
tps_p <- Virginia.ts %>%
  ggplot(aes(x = DATE)) +
  geom_line(aes(y = Y.Death, color = "Reported")) +
  geom_line(aes(y = tps_preds, color = "Fitted")) +
  scale_color_manual(
    values = c(Reported = "black", Fitted = "red")
  ) +
  labs(y = "Death count",
       title = "Reported vs truncated power spline fit") +
  guides(color = guide_legend(title = "Series"))

tps_p


```



## Natural spline regression model with 8 interior knots.

```{r}

library(splines)
n <- nrow(Virginia.ts)
t <- 1:n
ns_fit <- lm(Y.Death ~ ns(t, df = 6), data = Virginia.ts)
summary(ns_fit)
 Virginia.ts$ns_preds <- predict(ns_fit)

# Natural spline prediction and prediction intervals
h <- 14
t.new <- t[n] + (1:h)
ns_PI <- predict(ns_fit, newdata = data.frame(t = t.new), 
                 interval = "prediction", level = 0.95)

ns_PI <- as.data.frame(ns_PI) %>%
  mutate(DATE = ( Virginia.ts$DATE)[n] + 1:h)

# Plot of reported vs natural spline fit
ns <- Virginia.ts%>%
  ggplot(aes(x = DATE)) +
  geom_line(aes(y = Y.Death, color = "Reported")) +
  geom_line(aes(y = ns_preds, color = "Fitted")) +
  scale_color_manual(
    values = c(Reported = "black", Fitted = "red")
  ) +
  labs(y = "Number of deaths",
       title = "Reported vs natural spline regression fit") +
  guides(color = guide_legend(title = "Series")) +
  theme(legend.position = "bottom")

ns

# Plot of natural spline fit and its prediction intervals
ns_int <- ns + 
  geom_ribbon(
    mapping = aes(y = fit, 
                  ymin = lwr, 
                  ymax = upr,
                  fill = '95% Prediction Intervals'), 
    data = ns_PI, alpha = 0.2) + 
  geom_line(mapping = aes(y = fit, color = "Fitted"),
            data = ns_PI,
            key_glyph = "timeseries") +
  labs(title = "Natural spline regression fit and prediction intervals")  +
  guides(color = guide_legend(title = "Series"),
         fill =  guide_legend(title = "")) +
  theme(legend.position = "bottom")

ns_int



```



### Smoothing spline regression model with knots automatically selected by the “mgcv” package.

```{r}
if(!require('mgcv')) install.packages('mgcv')
library(mgcv)
```


```{r}
ss_fit <- gam(Y.Death ~ s(t, bs = "cr"), data = Virginia.ts)
summary(ss_fit)
Virginia.ts$ss_preds <- predict(ss_fit)

# Plot of reported vs smoothing spline fit
Virginia.ts %>%
  ggplot(aes(x = DATE)) +
  geom_line(aes(y = Y.Death, color = "Reported")) +
  geom_line(aes(y = ss_preds, color = "Fitted")) +
  scale_color_manual(
    values = c(Reported = "black", Fitted = "red")) +
  labs(y = "Number of deaths",
       title = "Reported vs smoothing spline fit") +
  guides(color = guide_legend(title = "Series"))  +
  theme(legend.position = "bottom")

```
